Sweat electrolytes and CFTR activity in the era of CFTR modulators

### 2024-09-06 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.qmd
CGM pre/post Trikafta responses

---
title: "Pre-/Post-Trikafta in CFRD"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(gtsummary)
library(arsenal)
library(rspiro)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r}
load("./Colleen Wood/Prepost Trikafta/Data_Cleaned/analysis_dataset.rdata")
# Baseline HbA1c, BMI group pre- and post-modulator
df <- df %>%
  group_by(`CFF ID`) %>%
  mutate(
    `Baseline HbA1c` = HbA1c[`Pre or Post` == "Pre"],
    `Baseline BMI` = `BMI Percentile`[`Pre or Post` == "Pre"],
    `Post BMI` = `BMI Percentile`[`Pre or Post` == "Post"]
  ) %>%
  ungroup()
df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf), right = F)
df$`Baseline BMI` <- cut(df$`Baseline BMI`, c(-Inf, 85, Inf), right = F)
df$`Post BMI` <- cut(df$`Post BMI`, c(-Inf, 85, Inf), right = F)
# Pre-modulator BMI percentile
df <- df %>%
  group_by(`CFF ID`) %>%
  mutate(`Baseline HbA1c` = HbA1c[`Pre or Post` == "Pre"]) %>%
  ungroup()
df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf), right = F)
# Race neutral equations
df$ppFEV1gl <- pctpred_GLIgl(
  age = df$Age, height = df$Height / 100, gender = df$Sex, FEV1 = df$FEV1
)
df$ppFVCgl <- pctpred_GLIgl(
  age = df$Age, height = df$Height / 100, gender = df$Sex, FVC = df$FEV1
)
# Columns formats
df$`Num. A1c Values` <- as.numeric(df$`Num. A1c Values`)
```

@@

```{r}
df %>%
  filter(`Pre or Post` == "Pre") %>%
  select(
    Age, `BMI Percentile`, Weight, Sex, ppFEV1, ppFEV1gl, ppFVC,
    ppFVCgl, `G-tube in past 12 months`, `Pancreatic Status`, Site,
    Genotype, `Age Group`, Race
  ) %>%
  tbl_summary(
    statistic = list(c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~ "{mean} ± {sd}"),
    missing_text = "Missing"
  )
```

@@

```{r}
# Table
df %>%
  select(
    `CFF ID`, `Pre or Post`, Site, `Num. A1c Values`, HbA1c, Time,
    Weight, `BMI Percentile`, ppFEV1, ppFEV1gl, ppFVC, ppFVCgl, TDD,
  ) %>%
  tbl_strata(
    strata = Site,
    .tbl_fun = ~ .x %>%
      tbl_summary(
        by = `Pre or Post`,
        type = list(
          c(
            "Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile",
            "ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl", "TDD"
          ) ~ "continuous2"
        ),
        statistic = list(
          c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~
            c("{mean} ± {sd}", "{min}, {max}"),
          c("Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile") ~
            c("{median} [{p25}, {p75}]", "{min}, {max}")
        ),
        include = -`CFF ID`, digits = list(everything() ~ 2),
        missing_text = "Missing"
      ) %>%
      add_p(
        test = list(
          c(
            "Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile",
            "TDD"
          ) ~ "paired.wilcox.test",
          c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~ "paired.t.test"
        ),
        group = `CFF ID`
      ) %>%
      separate_p_footnotes()
  )
# Plot
ggplot(df, aes(x = `Pre or Post`, y = HbA1c, group = `CFF ID`)) +
  geom_point(alpha = 0.2) +
  geom_line(alpha = 0.2) +
  theme_classic() +
  theme(axis.title.x = element_blank()) +
  facet_wrap(~Site)
```

@@

```{r}
cgm_vars <- c(cgm_vars, "TDD", "Long acting insulin dose")
cgm %>%
  select(`CFF ID`, `Pre or Post`, all_of(cgm_vars)) %>%
  tbl_summary(
    by = `Pre or Post`,
    type = list(everything() ~ "continuous2"),
    statistic = list(
      everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
    ),
    include = -`CFF ID`, digits = list(everything() ~ 2),
    missing_text = "Missing"
  ) %>%
  add_p(
    test = list(everything() ~ "paired.wilcox.test"),
    group = `CFF ID`
  ) %>%
  separate_p_footnotes()
```

@@

```{r}
cgm %>%
  select(
    `CFF ID`, `Pre or Post`, `Baseline HbA1c`, all_of(cgm_vars)
  ) %>%
  tbl_strata(
    strata = `Baseline HbA1c`,
    .tbl_fun = ~ .x %>%
      tbl_summary(
        by = `Pre or Post`,
        type = list(everything() ~ "continuous2"),
        statistic = list(
          everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
        ),
        include = -`CFF ID`, digits = list(everything() ~ 2),
        missing_text = "Missing"
      ) %>%
      add_p(
        test = list(everything() ~ "paired.wilcox.test"),
        group = `CFF ID`
      ) %>%
      separate_p_footnotes()
  )
```

@@

```{r}
cgm %>%
  select(
    `CFF ID`, `Pre or Post`, `Baseline BMI`, all_of(cgm_vars)
  ) %>%
  tbl_strata(
    strata = `Baseline BMI`,
    .tbl_fun = ~ .x %>%
      tbl_summary(
        by = `Pre or Post`,
        type = list(everything() ~ "continuous2"),
        statistic = list(
          everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
        ),
        include = -`CFF ID`, digits = list(everything() ~ 2),
        missing_text = "Missing"
      ) %>%
      add_p(
        test = list(everything() ~ "paired.wilcox.test"),
        group = `CFF ID`
      ) %>%
      separate_p_footnotes()
  )
```

@@

```{r}
cgm %>%
  select(
    `CFF ID`, `Pre or Post`, `Post BMI`, all_of(cgm_vars)
  ) %>%
  tbl_strata(
    strata = `Post BMI`,
    .tbl_fun = ~ .x %>%
      tbl_summary(
        by = `Pre or Post`,
        type = list(everything() ~ "continuous2"),
        statistic = list(
          everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
        ),
        include = -`CFF ID`, digits = list(everything() ~ 2),
        missing_text = "Missing"
      ) %>%
      add_p(
        test = list(everything() ~ "paired.wilcox.test"),
        group = `CFF ID`
      ) %>%
      separate_p_footnotes()
  )
```

@@

@@

@@

>>>>

---
title: "Pre-/Post-Trikafta in CFRD"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /home/tim/Documents/Miscellaneous/zotero.bib
csl: /home/tim/GitHub/styles/american-medical-association.csl
editor: source
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpmisc)
library(lubridate)
library(gtsummary)
library(rspiro)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood",
  "Linux" = "/home/tim/OneDrive/Vigers/CF/Colleen Wood"
)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| include: false
load("/home/tim/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/analysis_dataset.rdata")
# Ethnicity
chco_demo <- read.csv("/home/tim/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Raw/Data From CHCO/Registry/Copy of Earliest Trikafta Dates in Registry_updated 2.2.21 cc modified_cfrdyn.csv")
chco_demo$Hispanic <- NA
chco_demo$Hispanic[chco_demo$Hispanic.Latinx] <- "Yes"
chco_demo$Hispanic[!chco_demo$Hispanic.Latinx] <- "No"
chco_demo <- chco_demo[, c("CFF.ID", "Hispanic")]

montana_demo <- read.csv("/home/tim/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/montana_demo.csv")
montana_demo$Hispanic <- montana_demo$Is.Patient.of.Hispanic.Origin
montana_demo <- montana_demo[, c("CFF.ID", "Hispanic")]

uw_demo <- read.csv("/home/tim/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/uw_demo.csv")
uw_demo$Hispanic <- uw_demo$Is.Patient.of.Hispanic.Origin
uw_demo <- uw_demo[, c("CFF.ID", "Hispanic")]

eth <- do.call(rbind, list(chco_demo, montana_demo, uw_demo))
colnames(eth) <- c("CFF ID", "Hispanic")

df <- left_join(df, eth)
```

```{r}
# Baseline HbA1c, BMI group pre- and post-modulator, age group
df <- df %>%
  group_by(`CFF ID`) %>%
  mutate(
    `Baseline HbA1c` = HbA1c[`Pre or Post` == "Pre"],
    `Age group at baseline` = Age[`Pre or Post` == "Pre"],
    `Baseline BMI` = `BMI Percentile`[`Pre or Post` == "Pre"],
    `Post BMI` = `BMI Percentile`[`Pre or Post` == "Post"]
  ) %>%
  ungroup()
df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf),
  right = F
)
df$`Age group at baseline` <- cut(df$`Age group at baseline`,
  c(-Inf, 18, Inf),
  right = F
)
df$`Baseline BMI` <- cut(df$`Baseline BMI`, c(-Inf, 85, Inf), right = F)
df$`Post BMI` <- cut(df$`Post BMI`, c(-Inf, 85, Inf), right = F)
# Race neutral equations
df$ppFEV1gl <- pctpred_GLIgl(
  age = df$Age, height = df$Height / 100, gender = df$Sex, FEV1 = df$FEV1
)
df$ppFVCgl <- pctpred_GLIgl(
  age = df$Age, height = df$Height / 100, gender = df$Sex, FVC = df$FEV1
)
# TDD per kg
df$TDD <- df$TDD / df$Weight
# Columns formats
df$`Num. A1c Values` <- as.numeric(df$`Num. A1c Values`)
```

@@

```{r}
df %>%
  filter(`Pre or Post` == "Pre") %>%
  select(
    Age, `BMI Percentile`, Weight, Sex, ppFEV1, ppFEV1gl, ppFVC,
    ppFVCgl, `G-tube in past 12 months`, `Pancreatic Status`, Site,
    Genotype, `Age Group`, Race, Hispanic
  ) %>%
  tbl_summary(
    statistic = list(c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~
      "{mean} ± {sd}"),
    missing_text = "Missing"
  )
```

@@

```{r}
# Table
df %>%
  select(
    `CFF ID`, `Pre or Post`, Site, `Num. A1c Values`, HbA1c, Time,
    Weight, `BMI Percentile`, ppFEV1, ppFEV1gl, ppFVC, ppFVCgl, TDD,
  ) %>%
  tbl_strata(
    strata = Site,
    .tbl_fun = ~ .x %>%
      tbl_summary(
        by = `Pre or Post`,
        type = list(
          c(
            "Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile",
            "ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl", "TDD"
          ) ~ "continuous2"
        ),
        statistic = list(
          c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~
            c("{mean} ± {sd}", "{min}, {max}"),
          c("Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile") ~
            c("{median} [{p25}, {p75}]")
        ),
        include = -`CFF ID`, digits = list(everything() ~ 2),
        missing_text = "Missing"
      ) %>%
      add_p(
        test = list(
          c(
            "Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile",
            "TDD"
          ) ~ "paired.wilcox.test",
          c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~ "paired.t.test"
        ),
        group = `CFF ID`
      ) %>%
      separate_p_footnotes()
  )
# Plot
ggplot(df, aes(x = `Pre or Post`, y = HbA1c, group = `CFF ID`)) +
  geom_point(alpha = 0.2) +
  geom_line(alpha = 0.2) +
  theme_classic() +
  theme(axis.title.x = element_blank()) +
  facet_wrap(~Site)
```

### By age group at baseline

```{r}
# Table
df %>%
  select(
    `CFF ID`, `Pre or Post`, `Age group at baseline`, `Num. A1c Values`, HbA1c, Time,
    Weight, `BMI Percentile`, ppFEV1, ppFEV1gl, ppFVC, ppFVCgl, TDD,
  ) %>%
  tbl_strata(
    strata = `Age group at baseline`,
    .tbl_fun = ~ .x %>%
      tbl_summary(
        by = `Pre or Post`,
        type = list(
          c(
            "Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile",
            "ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl", "TDD"
          ) ~ "continuous2"
        ),
        statistic = list(
          c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~
            c("{mean} ± {sd}", "{min}, {max}"),
          c("Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile") ~
            c("{median} [{p25}, {p75}]", "{min}, {max}")
        ),
        include = -`CFF ID`, digits = list(everything() ~ 2),
        missing_text = "Missing"
      ) %>%
      add_p(
        test = list(
          c(
            "Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile",
            "TDD"
          ) ~ "paired.wilcox.test",
          c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~ "paired.t.test"
        ),
        group = `CFF ID`
      ) %>%
      separate_p_footnotes()
  )
# Plot
ggplot(df, aes(x = `Pre or Post`, y = HbA1c, group = `CFF ID`)) +
  geom_point(alpha = 0.2) +
  geom_line(alpha = 0.2) +
  theme_classic() +
  theme(axis.title.x = element_blank()) +
  facet_wrap(~`Age group at baseline`)
```

@@

```{r}
cgm_vars <- c(cgm_vars, "TDD", "Long acting insulin dose")
cgm %>%
  select(`CFF ID`, `Pre or Post`, all_of(cgm_vars)) %>%
  tbl_summary(
    by = `Pre or Post`,
    type = list(everything() ~ "continuous2"),
    statistic = list(
      everything() ~ c("{median} [{p25}, {p75}]")
    ),
    include = -`CFF ID`, digits = list(everything() ~ 2),
    missing_text = "Missing"
  ) %>%
  add_p(
    test = list(everything() ~ "paired.wilcox.test"),
    group = `CFF ID`
  ) %>%
  separate_p_footnotes()
```

#### By age group at baseline

```{r}
cgm %>%
  select(
    `CFF ID`, `Pre or Post`, `Age group at baseline`, all_of(cgm_vars)
  ) %>%
  tbl_strata(
    strata = `Age group at baseline`,
    .tbl_fun = ~ .x %>%
      tbl_summary(
        by = `Pre or Post`,
        type = list(everything() ~ "continuous2"),
        statistic = list(
          everything() ~ c("{median} [{p25}, {p75}]")
        ),
        include = -`CFF ID`, digits = list(everything() ~ 2),
        missing_text = "Missing"
      ) %>%
      add_p(
        test = list(everything() ~ "paired.wilcox.test"),
        group = `CFF ID`
      ) %>%
      separate_p_footnotes()
  )
```

@@

```{r}
cgm %>%
  select(
    `CFF ID`, `Pre or Post`, `Baseline HbA1c`, all_of(cgm_vars)
  ) %>%
  tbl_strata(
    strata = `Baseline HbA1c`,
    .tbl_fun = ~ .x %>%
      tbl_summary(
        by = `Pre or Post`,
        type = list(everything() ~ "continuous2"),
        statistic = list(
          everything() ~ c("{median} [{p25}, {p75}]")
        ),
        include = -`CFF ID`, digits = list(everything() ~ 2),
        missing_text = "Missing"
      ) %>%
      add_p(
        test = list(everything() ~ "paired.wilcox.test"),
        group = `CFF ID`
      ) %>%
      separate_p_footnotes()
  )
```

@@

```{r}
cgm %>%
  select(
    `CFF ID`, `Pre or Post`, `Baseline BMI`, all_of(cgm_vars)
  ) %>%
  tbl_strata(
    strata = `Baseline BMI`,
    .tbl_fun = ~ .x %>%
      tbl_summary(
        by = `Pre or Post`,
        type = list(everything() ~ "continuous2"),
        statistic = list(
          everything() ~ c("{median} [{p25}, {p75}]")
        ),
        include = -`CFF ID`, digits = list(everything() ~ 2),
        missing_text = "Missing"
      ) %>%
      add_p(
        test = list(everything() ~ "paired.wilcox.test"),
        group = `CFF ID`
      ) %>%
      separate_p_footnotes()
  )
```

@@

```{r}
cgm %>%
  select(
    `CFF ID`, `Pre or Post`, `Post BMI`, all_of(cgm_vars)
  ) %>%
  tbl_strata(
    strata = `Post BMI`,
    .tbl_fun = ~ .x %>%
      tbl_summary(
        by = `Pre or Post`,
        type = list(everything() ~ "continuous2"),
        statistic = list(
          everything() ~ c("{median} [{p25}, {p75}]")
        ),
        include = -`CFF ID`, digits = list(everything() ~ 2),
        missing_text = "Missing"
      ) %>%
      add_p(
        test = list(everything() ~ "paired.wilcox.test"),
        group = `CFF ID`
      ) %>%
      separate_p_footnotes()
  )
```

@@

#### By age group at baseline

```{r}
cgm %>%
  select(
    `CFF ID`, `Pre or Post`, `Age group at baseline`, all_of(day_cgm)
  ) %>%
  tbl_strata(
    strata = `Age group at baseline`,
    .tbl_fun = ~ .x %>%
      tbl_summary(
        by = `Pre or Post`,
        type = list(everything() ~ "continuous2"),
        statistic = list(
          everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
        ),
        include = -`CFF ID`, digits = list(everything() ~ 2),
        missing_text = "Missing"
      ) %>%
      add_p(
        test = list(everything() ~ "paired.wilcox.test"),
        group = `CFF ID`
      ) %>%
      separate_p_footnotes()
  )
```

@@

#### By age group at baseline

```{r}
cgm %>%
  select(
    `CFF ID`, `Pre or Post`, `Age group at baseline`, all_of(night_cgm)
  ) %>%
  tbl_strata(
    strata = `Age group at baseline`,
    .tbl_fun = ~ .x %>%
      tbl_summary(
        by = `Pre or Post`,
        type = list(everything() ~ "continuous2"),
        statistic = list(
          everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
        ),
        include = -`CFF ID`, digits = list(everything() ~ 2),
        missing_text = "Missing"
      ) %>%
      add_p(
        test = list(everything() ~ "paired.wilcox.test"),
        group = `CFF ID`
      ) %>%
      separate_p_footnotes()
  )
```

@@

# Additional analyses

## Longitudinal HbA1c

The vast majority of participants have 2 A1c values pre-ETI and 2 after.

```{r}
df %>%
  mutate(`Num. A1c Values` = factor(`Num. A1c Values`)) %>%
  select(`Pre or Post`, `Num. A1c Values`) %>%
  tbl_summary(by = `Pre or Post`)
```

## Change in HbA1c and change in FEV1

```{r}
change <- df %>%
  group_by(`CFF ID`) %>%
  summarise(
    `Baseline HbA1c` = HbA1c[`Pre or Post` == "Pre"],
    `Baseline ppFEV1gl` = ppFEV1gl[`Pre or Post` == "Pre"],
    `Post ppFEV1gl` = ppFEV1gl[`Pre or Post` == "Post"],
    `Change in HbA1c` =
      HbA1c[`Pre or Post` == "Post"] - HbA1c[`Pre or Post` == "Pre"],
    `Change in FEV1` =
      ppFEV1gl[`Pre or Post` == "Post"] - ppFEV1gl[`Pre or Post` == "Pre"],
    `% Change in FEV1` =
      ((ppFEV1gl[`Pre or Post` == "Post"] - ppFEV1gl[`Pre or Post` == "Pre"]) /
        ppFEV1gl[`Pre or Post` == "Pre"]) * 100
  )
change$`HbA1c group` <- cut(change$`Change in HbA1c`,
  breaks = c(-Inf, -0.5, 0.5, Inf), right = F,
  labels = c("Decreased", "No change", "Increased")
)
ggplot(change, aes(x = `Change in FEV1`, y = `Change in HbA1c`)) +
  geom_point() +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2", "p", "n"))) +
  theme_bw()
ggplot(change, aes(x = `% Change in FEV1`, y = `Change in HbA1c`)) +
  geom_point() +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2", "p", "n"))) +
  theme_bw()
```

## Change in FEV1 by HbA1c group

```{r}
df <- df %>%
  group_by(`CFF ID`) %>%
  mutate(
    `Change in HbA1c` =
      HbA1c[`Pre or Post` == "Post"] - HbA1c[`Pre or Post` == "Pre"]
  )
df$`HbA1c group` <- cut(df$`Change in HbA1c`,
  breaks = c(-Inf, -0.5, 0.5, Inf), right = F,
  labels = c("HbA1c decreased", "HbA1 no change", "HbA1c increased")
)
ggplot(df, aes(x = `Pre or Post`, y = ppFEV1gl, group = `CFF ID`)) +
  geom_point(alpha = 0.2) +
  geom_line(alpha = 0.2) +
  theme_classic() +
  theme(axis.title.x = element_blank()) +
  facet_wrap(~`HbA1c group`)
```

### 2024-09-06 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
CGM pre/post Trikafta responses

```{r setup}
#| include: false
library(tidyverse)
library(readxl)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Test",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Test",
  "Linux" = "/home/tim/OneDrive/Vigers/CF/Edith Zemanick/Sweat Test"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}

```

>>>>

```{r setup}
#| include: false
library(dplyr)
library(tidyr)
library(readxl)
library(gtsummary)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Test",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Test",
  "Linux" = "/home/tim/OneDrive/Vigers/CF/Edith Zemanick/Sweat Test"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
# Import data
chco <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv",
  na.strings = ""
)
njh <- read_excel("./Data_Raw/SweatVariability2025_CFgenotypes 2024-08-21.xlsx")
chec <- read.csv("./Data_Raw/CHEC sweat sodium and chloride.csv")
# Numeric variables
num_vars <- c(
  "Na", "K", "Cl", "Amount_Left", "Amount_Right", "Na_Left", "Na_Right",
  "K_Left", "K_Right", "Cl_Left", "Cl_Right"
)
chco[, num_vars] <- lapply(chco[, num_vars], function(c) {
  suppressWarnings(as.numeric(c))
})
# Fix diagnoses/exclusion data for CHCO (based on Angela's review)
diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) | diagnoses$Code == "Exclude"]
exclude <- exclude[!is.na(exclude)]
chco <- chco[!(chco$PatientID %in% exclude), ]
chco$Diagnosis <- factor(
  diagnoses$Diagnosis[match(
    chco$PatientID,
    diagnoses$PatientID
  )],
  levels = c("CFTR-related disorder", "CRMS", "Cystic Fibrosis", "Non-CF"),
  labels = c("CFTR-RD or CRMS", "CFTR-RD or CRMS", "Cystic Fibrosis", "Non-CF")
)
```

```{r}
chco %>%
  group_by(PatientID) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(-PatientID, -Genotypes1, -Genotypes2, -T_Tract, -Notes) %>%
  tbl_summary(by = Diagnosis, missing_text = "(Missing)")%>%
  add_overall()
```
